---
title: "Fluid Overload Creation"
author: "Tianyi Zhang"
date: "2023-03-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("~/Library/CloudStorage/OneDrive-UniversityofGeorgia/ShiyuanD/DataProcessing")
library(readr)
library(dplyr)
suppressMessages(library("readxl"))
flowsheets_alt<- read_csv("flowsheets_w_ht_wt.csv",
                      col_types = cols(MEAS_VALUE = col_character()))
ICU_stay_time=read_csv("ICU Stay Time.csv")
#patient=(ICU_stay_time%>%filter(end-start>2))$PATIENT_DEID
#patients=(data_fluid.overload%>%filter(!is.na(fluid_overload)))$PATIENT_DEID
#patients%in%patient
flowsheets_alt = left_join(flowsheets_alt,ICU_stay_time)

data_my=read.csv("fluid_all.csv",header=T)
data_xl=read_excel(path="master_table_0608.xlsx",sheet="Sheet1",
                   range="A1:AY992",col_names=T,na="NA")
id_remove=data_my$PATIENT_DEID[!data_my$PATIENT_DEID%in%data_xl$PATIENT_DEID]
```

# Fluid Overload

## Fluid Overload based on weight

Definition: a patient experienced fluid overload when the maximum weight measure in the ICU is at least 110% of the admission weight/ the very first weight measure available. 

Note: admission weight is at admission, not from the start of ICU.

Weight info: 
- In the flowsheets_w_ht_wt.csv file (object name: flowsheets_alt)
- unit: oz; need to tranform to kg (formula oz to kg: divide the mass value by 35.274)
- we only have weight measurement for 992 patients
- only 947 patients have weight records when they were the ICU
(* only 865 records in the flowsheets file updated on 2021_11_24????)

```{r}
require(tidyverse)
weight_ICU =  flowsheets_alt%>%filter(str_detect(FLO_MEAS_NAME,"WEIGHT"))%>%
  mutate(weight = as.numeric(MEAS_VALUE)) %>%
  dplyr::select(PATIENT_DEID,RECORDED_TIME_DEID,weight,start,end) %>%
  mutate(weight_kg = weight/35.274)%>%
  filter(weight_kg > 1)%>%
  filter(#(difftime(RECORDED_TIME_DEID,start,units="hours")>48),
         (difftime(end,RECORDED_TIME_DEID,units="hours")>=0),
         (difftime(RECORDED_TIME_DEID,start,units="hours")<=72)
         )

weight_max_ICU = weight_ICU%>%
  group_by(PATIENT_DEID) %>% 
  summarise(max_ICU_weight = max(weight_kg))

weight_admission = flowsheets_alt%>%filter(str_detect(FLO_MEAS_NAME,"WEIGHT"))%>%
  mutate(weight = as.numeric(MEAS_VALUE)) %>%
  dplyr::select(PATIENT_DEID,RECORDED_TIME_DEID,weight)%>%
  mutate(weight_kg = weight/35.274)%>%
  filter(weight_kg > 1)%>%
  group_by(PATIENT_DEID) %>% 
  summarise(earliest_record = min(RECORDED_TIME_DEID), 
            admission_weight_kg = weight_kg[RECORDED_TIME_DEID==earliest_record])%>%
  dplyr::select(PATIENT_DEID,admission_weight_kg)%>%
  filter(!PATIENT_DEID%in%id_remove)

length(unique(weight_ICU$PATIENT_DEID))
length(unique(weight_admission$PATIENT_DEID))


```


Fluid overload:

Among all patients, 101 patients have fluid overload, 764 did not, and we don't have weight records to determine the state of fluid for the rest.

```{r}
fluid_wt = full_join(weight_admission,weight_max_ICU, by ="PATIENT_DEID" )%>%
  mutate(fluid_overload_wt = max_ICU_weight/admission_weight_kg)%>%
  dplyr::select(PATIENT_DEID,fluid_overload_wt)
```


For the 991 patients:
- 101 had fluid overload
- 760 did not have fluid overload
- 130 have no weight records in ICU to determine fluid status.


## Fluid Overload based on volumn


From the I_O_data file

Instruction:
- net fluid balance after day 1, day 2, etc. in mL (a positive or negative volume in milliliters)

- presence or absence of fluid overload after day 1, day 2, etc. (0 = absence, 1 = presence, presence being defined by an amount > 10% of their admission weight in milliliters). 


Problem: unit mL vs 0.1\*kg
Check with Andrea (04/03): unit L vs 0.1\* kg

Another issue: daily net or cumulative. Multiple daily net values.


```{r}
I_O_data = read_csv("I_O_data.csv")

fluid_balance = I_O_data%>%filter(!is.na(IO_ID))%>%dplyr::select(PATIENT_DEID,RECORDED_DATE_DEID,DAILY_NET)%>%distinct()
fluid_balance = left_join(fluid_balance,ICU_stay_time)%>%
  filter(RECORDED_DATE_DEID>=as.Date(start),
       #RECORDED_DATE_DEID>=as.Date(start)+2, 
         RECORDED_DATE_DEID<=as.Date(end),
         RECORDED_DATE_DEID<=as.Date(start)+3
         )%>%
  arrange(PATIENT_DEID,RECORDED_DATE_DEID)%>%group_by(PATIENT_DEID)%>%
  mutate(cum_fluid = cumsum(DAILY_NET))%>%group_by(PATIENT_DEID,RECORDED_DATE_DEID)%>%
  mutate(cum_fluid_daily_total = max(cum_fluid),
            start = start)%>%distinct()

max_fluid_in_ICU = fluid_balance%>%
  group_by(PATIENT_DEID)%>%
  summarise(max_fluid_ICU = max(cum_fluid_daily_total,na.rm = T))

fluid_volume = left_join(max_fluid_in_ICU, weight_admission)%>%
  mutate(fluid_overload_volume = 1+max_fluid_ICU/(admission_weight_kg*1000))%>%
  dplyr::select(PATIENT_DEID,fluid_overload_volume)
```

```{r}
fluid_overload=full_join(fluid_wt,fluid_volume, by ="PATIENT_DEID")%>%
  filter(!PATIENT_DEID%in%id_remove)
fluid_overload ##NA?
```




114 no, 872 yes, 8 unknown (comparing mL to 0.1 weight in kg)

969 no, 17 yes, 8 unknown.

875 no, 105 yes, 8 unknown. (cumulative balance)




## combine the two conditions

rule: 
- TRUE = TRUE and whatever (FALSE/missing)
- FALSE = FALSE and FALSE
- unknown = the rest combination


```{r}
fluid_overload1.1%>%
  filter(fluid_overload_volume>=1.1)
fluid_overload1.07%>%
  filter(fluid_overload_volume>=1.07)

fluid_all = full_join(fluid_wt,fluid_volume1.1,fluid_volume1.07)%>%
  mutate(fluid_overload_volume_nu = as.numeric(fluid_overload_volume))%>%
  mutate(fluid_overload_wt_nu =as.numeric(fluid_overload_volume))%>%
  mutate(fluid_overload = fluid_overload_volume|fluid_overload_wt)%>%
  mutate(fluid_overload = ifelse(fluid_overload>=1,1, fluid_overload))%>%
  dplyr::select(PATIENT_DEID,fluid_overload, fluid_overload_volume,fluid_overload_wt)


#master_sheet = left_join(master_sheet,fluid_all)


table(fluid_all$fluid_overload, useNA = "ifany")

prop.table(table(fluid_all$fluid_overload))

saveRDS(fluid_all,file='fo.rds')
```

